On Pure Exploration, Thresholding Bandits, and Kullback-Leibler Divergence

### 2017-09-18 Thesis/Draft.Rmd
final

---
title: "On Pure Exploration, Thresholding Bandits, and Kullback-Leibler Divergence"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \usepackage{hyperref}
   - \usepackage[lined,boxed]{algorithm2e}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{corollary}{Corollary}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   - \newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    toc: true
    toc_depth: 2
    #latex_engine: xelatex
fontsize: 11pt
urlcolor: blue
---

@@

"In general, we look for a new law by the following process: First we guess it. Then we – now don't laugh, that's really true. Then we compute the consequences of the guess to see what, if this is right, if this law that we guessed is right, to see what it would imply. And then we compare the computation results to nature, or we say compare to experiment or experience, compare it directly with observations to see if it works. If it disagrees with experiment, it's wrong. In that simple statement is the key to science. It doesn't make any difference how beautiful your guess is, it doesn't make any difference how smart you are, who made the guess, or what his name is. If it disagrees with experiment, it's wrong. That's all there is to it."

- Richard Feynman in *Richard Feynman Messenger Lectures at Cornell University: The Character of Physical Law. Chapter 7, Seeking New Laws*.

@@

@@

This makes sequential sampling especially relevant in situations where there is an inherent cost to obtaining samples, or where only a limited amount of samples can be collected. The classical example is that of clinical trials in which a researcher tests two drugs on patients to observe which one performs better. Not only is it difficult to gather enough patients, but one would also like to adapt the sampling such that as soon as possible more patients receive the better drug and are cured. As noted in Cappé et al. (2013), this example dates back to Thompson (1933, 1935). Such problems are known as multi-armed bandit problems in which the sampling is adapted so as to minimize the cumulative regret: In the case of the drug tests, one would like to sample the two drugs in a way to find the best drug as soon as possible and then to test this drug more and more often and more often than the worse drug. This would lead to cumulatively more healthy patients than testing both drugs uniformly; it would minimize the regret from treating patients with an ineffective drug.

Instead of drugs, one can consider two or more unknown distributions (which we call arms in the multi-armed bandit problem in analogy with the one-armed bandit slot machine). These distributions have different means, and one can draw samples sequentially from the distributions in order to estimate their means. To minimize the cumulative regret measure, one would like to play the distribution with the highest mean as much as possible (*Exploitation*). At the same time, however, one has to sample all arms at least a little, so as to initially find the best arm (*Exploration*). Any algorithm thus has to handle a trade-off between exploration and exploitation to minimize the regret. How can we organize the sampling in order to achieve this task?

While Gittins (1979) presents an Bayesian-optimal solution to the cumulative regret problem, the Upper-Confidence Bound (UCB) algorithms pioneered by Agrawal (1995) and Auer, Cesa-Bianchi and Fischer (2002) have lead to a blooming literature on multi-armed bandit problems that has been picked up to solve modern problems in website and advertising optimization, where advertisers try to allocate their advertising optimally so as to reduce regret from showing poorly converting advertisements. Several algorithms like the KL-UCB algorithm based on Kullback-Leibler divergence upper confidence intervals (Cappé et al., 2013) or the Bayes-UCB algorithm (Kaufmann et al., 2012)  have followed and been shown to be optimal as they reach the lower bound introduced in Lai and Robbins (1985) and Burnetas and Katehakis (1996).

@@

Similar algorithms based on indexes can be used on related problems. Instead of minimizing cumulative regret, it can be of interest to find the best arm, that is, the distribution with the largest mean -- with as few samples as possible and with confidence as large as possible. This is particularly useful when there is not a cost from testing arms with lower mean, but when drawing any sample is costly. Thus one needs to explore the distributions as efficiently as possible. Exploitation is not necessary, and the setting is thus referred to as *Pure Exploration*. The need for strategies other than uniform sampling arises: One does not want to continue to sample distributions that one is confident in. This setting is related to the problem faced in Bayesian Optimization: Training large neural networks is costly in terms of time and computation, and so the parameter space needs to be explored with as few samples as possible. (Zhang et al., 2017; Brochu et al., 2010). The problem of best arm identification is also very related to statistical hypothesis testing as noted by Chernoff (1958) since one has to make a decision about how the means of distributions relate to each other. The main difference being that the bandit setting includes the testing procedure in the sampling process.

Given the focus on exploration, the pure exploration problem is usually described with a loss function different from the one in the cumulative regret settings. Audibert et al. (2010) introduce a soft measure called *simple regret* based on which they describe UCB policies. Other authors investigate binary measures that are equal to 1 when the best arm has not been identified (Russo, 2016; Garivier and Kaufmann, 2016). These binary measures correspond to the probability of the identified arm being the best, and thus inherently related to hypothesis testing. The problem of best arm identification can readily be extended to the so called Top-$m$ problem of identifying the best $m$ of $K$ arms (Kaufmann and Kalyanakrishnan, 2013; Kaufmann et al. 2016).

In this thesis, we consider the thresholding bandit problem. While the name is reminiscient of the cumulative regret multi-armed bandit, the problem is more similar to the top-$m$ problem. But instead of identifying the best $m$ out of $K$ arms, one aims to find all distributions out of $K$ that have a mean at least as large as a fixed threshold. Given that an implicit threshold arises in the top-$m$ problem between the $m$-th best and the $(m+1)$-th best arm (Kaufmann et al., 2016), the two problems share many characteristics. We discuss this in more detail in \autoref{chap:KullbackLeiblerPureExploration}. In the form described in Locatelli et al. (2016), algorithms in the thresholding bandit problem are evaluated against a binary loss function penalizing whenever an arm has not been classified correctly as being above or below the threshold, and thus focus on exploration similar to algorithms employed for best arm identification. Indeed, Locatelli et al. (2016) show how their thresholding algorithm can be used for best arm identification when the mean of the best arm is known upfront.

@@

The app will then display a "Good service" or "This place is trendy" description for the restaurant if $\mu_{\text{service}}$ and $\mu_{\text{trendy}}$ are estimated to be above a certain threshold, for example $\tau = 0.5$. That is, more than half of the people would describe the place as having good service, or as trendy.

To make the app as useful as possible to other users, this information should be shown as quickly as possible. But even though the app has thousands of users, not that many people visit a specific restaurant, and even less people report this back to the app. Thus the company only has a limited budget of $T$ potential users from which to collect samples $X_{\text{service}}$ and $X_{\text{trendy}}$ from. How can the company ask these $T$ questions to be as confident as possible in its classification, that is, in the information it displays?

A straightforward way of sampling users would be to ask the two questions in a round-robin fashion: Ask for "Service", then for "Trendy", then for "Service", ... Given 100 users, 50 users give feedback on the service level, 50 users report on the trendiness level. This however is unlikely to be optimal with respect to the confidence after 100 samples. Given that the company asks the users sequentially, there is the option to adapt the sampling strategy over time: If the first ten users report $\sum_{s=1}^{10}X_{\text{service},s}=10$, while $\sum_{s=1}^{10}X_{\text{trendy},s}=5$, then the company might already be quite confident that $\mu_{\text{service}} \geq \tau = 0.5$, while this is less clear for $\mu_{\text{trendy}}$. Instead of continuously sampling both questions, it might be more useful to collect more feedback on the trendiness level so as to figure out its position relative to the threshold. After all, the company would like to use the 100 samples to maximize its confidence in both displayed tips.

Finding the optimal sequential sampling strategy to classify distributions against a threshold with as much confidence as possible after a limited budget of samples is the problem described above, and the problem described by the thresholding bandit problem which we discuss for the remainder of this thesis. In the following section, we define the general setting of the thresholding bandit problem.

@@

Consider a setting with a set of $\mathbb{A}=\{1, ..., K\}$ unknown probability distributions $\nu_i, i \in \mathbb{A}$. To stick with the idea of multi-armed bandits, we call the distributions arms, and observations from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, she would like to estimate the means $\mu_1, ..., \mu_K$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from the different arms, where $X_{i} \sim \nu_i$. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which she would like to observe the reward $X_{I_t,t}$. Often, adaptive sampling strategies makes the decision of which arm to pull by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then is $B_i(t)$ a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indexes can be called strategy (Kaufmann et al., 2012) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

@@

The thresholding problem for the learner is now to use a budget of overall $T$ samples (or rounds) to explore the different arms. After this exploration phase, the learner returns a set of arms that she believes to be over the threshold. When evaluating the classification of the learner, arms with mean inside of the epsilon interval $\tau \pm \epsilon$ will not be considered. The learner only has to correctly distinguish the means from the threshold up to the precision given by $\epsilon$. Thus, the algorithm is not evaluated against $\mathcal{S}_{\tau}$ and $\mathcal{S}^C_{\tau}$, but needs to distinguish $\mathcal{S}_{\tau+\epsilon}$ from $\mathcal{S}^C_{\tau-\epsilon}$.

The learner's classification is returned as the set $\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) \subset \mathbb{A}$ after $T$ rounds; that is, the set of arms estimated to be over the threshold. The classification is evaluated against the loss given as (Locatelli et al., 2016):

@@

Algorithms for this thresholding problem can be compared based on their expected loss. Let $\mathbb{E}$ be the expectation with regard to the samples $X_{i,1}, ..., X_{i,T}$ collected by the algorithm. The expected loss is then given as

@@

which corresponds to the probability of making at least one wrong classification. Both the probability of not identifying an arm that is actually over the threshold, as well as the probability of accepting an arm as over the threshold when it is not, are evaluated. The oracle strategy would evaluate all arms correctly with an error probability of $0$ (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit problem setting which we will use throughout the thesis. The setting can be extended by different assumptions to make certain algorithms or bounds feasible. For that reason, Locatelli et al. (2016) assume that the distributions $\nu_i, i \in \mathbb{A}$ are $R$-sub-Gaussian. Under this assumption, the authors are able to prove a lower bound on the largest probability of error of the best algorithm in the setting we have just introduced, as well as an upper bound on the performance of their APT algorithm which we will introduce in the following chapter. As in their paper, we define $R$-sub-Gaussian and assume it in the description of the their algorithm and bounds:

@@

As noted by Locatelli et al. (2016), this ensures that the mean $\mu_i$ of the distributions $\nu_i$ considered in our problem are finite. $R$-sub-Gaussian distributions offer concentration inequalities for the concentration of the empirical mean around the true mean, which makes it possible to deduce the bounds presented by the authors and repeated below. The family of $R$-sub-Gaussian distributions encompasses Gaussian distributions of variance $R^2$ for $R \in \mathbb{R}$, as well as all bounded distributions, with the important case of Bernoulli distributions.

Having described the thresholding bandit setting as well as the assumption of $R$-sub-Gaussianity, we can now proceed to describe the *Anytime Parameter-free Thresholding* (APT) algorithm proposed in Locatelli et al. (2016) for the thresholding bandit problem.

@@

\IncMargin{1em}
\begin{algorithm}
\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$}
\Output{$\hat{S}_{\tau} = \{i \in \{1,...,K\}, \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{i \in A} B_i^{APT}(t) = \arg \min_{i \in A} \sqrt{T_i(t)} (|\hat{\mu}_i(t) - \tau|+\epsilon)$}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{Anytime Parameter-free Thresholding Algorithm (Locatelli et al., 2016)}\label{alg:APT}
\end{algorithm}
\DecMargin{1em}

Just like the general index policies mentioned in \autoref{sec:Setup}, we see that the APT algorithm adapts its strategy using an index $B_i(t+1)$ for each arm that is based entirely on the observations observed until time $t$. Given the arm with the smallest index is pulled at round $t+1$, we observe that arms that are estimated to be further away from the threshold $\tau$ as measured by the absolute distance to their empirical mean $\hat{\mu}_i$, $\hat{\Delta}_i(t)$, are pulled less often. Intuitively, it is easier to classify arms as below or above $\tau$ when their mean is far away from the threshold. In contrast to the uniform sampling strategy, APT will gather more evidence for arms which appear to be inherently more difficult to classify. Likewise, APT tends to less often pull arms which it already pulled considerably often as measured by $T_i(t)$. If arm $i$ has been pulled more often than arm $j$, then we should in general be more confident in its estimated mean and the corresponding distance to the threshold. If we are confident in $i$, then it is better to explore other arms by drawing an observation from arm $j$. By adapting its sampling strategy in this way during the sequential sampling process, the APT algorithm minimizes the probability of error $\mathbb{E}[\mathcal{L}(T)]$ after $T$ samples have been observed.

The previous paragraph describes how arms close to the threshold are intuitively more difficult to classify. In the index employed by the APT algorithm, the distance between an arm and the threshold is measured by the absolute distance $\hat{\Delta}_i(t)$ between the (estimated) mean of the arm and the threshold $\tau$ (and the epsilon interval). When the APT algorithm realizes that an arm has a mean close to the threshold, then this arm requires more attention than an arm far away from the threshold; the confidence in the empirical mean has to be larger. Or the empirical mean has to be better concentrated around the true mean. Otherwise it is easier to make a mistake on this arm. To gain this confidence, the arm requires more samples, and thus a larger share of the overall $T$ samples available. With this intuition, it makes sense to measure the complexity of a thresholding problem: If we knew upfront how complex a given set of arms is, we would know whether to expect a small or large probability of error given our budget $T$. Also, for example in the best arm identification problem, there exist algorithms which take the complexity of a problem as input. Based on the complexity they then adapt their sampling strategy. See for example the UCB-E algorithm in Audibert et al. (2010) for the best arm identification problem, or the KL-LUCB-E algorithm in Kaufmann and Kalyanakrishnan (2013) as extension to the top-$m$ problem.

While the APT algorithm does not require the complexity to be known upfront (which would be problematic as it is not known upfront), it is necessary to define the complexity to deduce the upper bound on its probability of error. Similar to the distance measure used in the algorithm's index, we define the gap $\Delta_i^{\tau,\epsilon}$ between arm $i$ and the threshold $\tau \pm \epsilon$ as

@@

Note that the complexity of the problem increases in the number of arms $K$ and decreases in the gap $\Delta_i^{\tau,\epsilon}$. Consequently, when we require less precision, that is, we increase $\epsilon$, the hardness decreases. Furthermore, we see that each arm contributes its individual complexity, $h_i^{\tau,\epsilon} = \frac{1}{(\Delta_i^{\tau,\epsilon})^2}$ linearly into the problem. Thus, if the complexity $H$ is a good measure of the actual problem complexity, and we knew upfront which arm has which mean, we could allocate the "optimal" share of the budget, $T \cdot h_i /H$ to arm $i$. However, since we do not know neither $\mu_i$ nor $H$ upfront, the best we can do is to adapt our allocation of samples more and more towards this optimum as we observe samples.

At this point, one can question whether the measure of complexity $H$ proposed in Locatelli et al. (2016) is a good measure of the "true" problem complexity. In the following \autoref{sec:UpperBoundLocatelli}, we will show the upper bound on the probability of error for the APT algorithm and see how the complexity $H$ arises naturally for the APT algorithm. In \autoref{sec:LowerBoundLocatelli} we then show the lower bound presented in Locatelli et al. (2016) and that the APT algorithm is optimal in that it matches the lower bound. We discuss the results and the APT algorithm further in \autoref{sec:DiscussionLocatelli}.

## Upper Bound for the APT Algorithm \label{sec:UpperBoundLocatelli}

@@

This bound is non-trivial for much smaller values of $T$, and we will use it in the following.

>>>>

---
header-includes:
   #- \documentclass[a4paper,oneside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \usepackage{hyperref}
   - \usepackage[lined,boxed]{algorithm2e}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{corollary}{Corollary}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   - \newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    #toc: true
    toc_depth: 2
    fig_caption: yes
    #latex_engine: xelatex
#documentclass: article
#classoption: oneside
#geometry: left=2.5cm, right=2cm
fontsize: 11pt
urlcolor: black
linkcolor: black
---

\pagenumbering{gobble}

```{r child = 'titlepage.Rmd'}
```

@@

"In general, we look for a new law by the following process: First we guess it. Then we – now don't laugh, that's really true. Then we compute the consequences of the guess to see what, if this is right, if this law that we guessed is right, to see what it would imply. And then we compare the computation results to nature, or we say compare to experiment or experience, compare it directly with observations to see if it works. If it disagrees with experiment, it's wrong. In that simple statement is the key to science. It doesn't make any difference how beautiful your guess is, it doesn't make any difference how smart you are, who made the guess, or what his name is. If it disagrees with experiment, it's wrong. That's all there is to it." \vspace*{1.5\baselineskip}

Richard Feynman in *Richard Feynman Messenger Lectures at Cornell University: The Character of Physical Law. Chapter 7, Seeking New Laws*.

@@

\Large{\bf Abstract\\}

\vspace*{1\baselineskip}

\normalsize
In this thesis we present a new sequential sampling strategy for the fixed-budget thresholding bandit problem described in Locatelli et al. (2016). We draw from the literature on pure exploration bandit problems to show the similarity of the thresholding bandit to (sequential) hypothesis testing. In particular, we derive a lower bound on how fast the error probability in a one-armed thresholding bandit problem can decrease at best, and show that this corresponds to the Chernoff-Stein Lemma. This motivates the use of the Kullback-Leibler divergence in the index of our new algorithm. In simulations, our strategy is shown to outperform existing strategies. Using website analytics data from an online shop, we use time series cross-validation to create data sets for offline testing of adaptive sampling strategies; time dependency in the data is thus preserved. Our experiments show that the adaptive sampling strategies are disrupted by the dependency structure in the samples.

\setcounter{tocdepth}{2}
\tableofcontents

\newpage
\pagenumbering{arabic}

@@

This makes sequential sampling especially relevant in situations where there is an inherent cost to obtaining samples, or where only a limited amount of samples can be collected. The classical example is that of clinical trials in which a researcher tests two drugs on patients to observe which one performs better. Not only is it difficult to gather enough patients, but one would also like to adapt the sampling such that more patients receive the better drug as soon as possible, and are cured. As noted in Cappé et al. (2013), this example dates back to Thompson (1933, 1935). Such problems are known as multi-armed bandit problems in which the sampling is adapted so as to minimize the cumulative regret: In the case of the drug tests, one would like to sample the two drugs in a way that quickly finds the best drug and then tests this drug more and more often, and more often than the worse drug. This would lead to cumulatively more healthy patients than testing both drugs uniformly; it would minimize the regret from treating patients with an ineffective drug.

Instead of drugs, one can consider two or more unknown distributions (which we call arms in the multi-armed bandit problem in analogy with the one-armed bandit slot machine). These distributions have different means, and one can draw samples sequentially from the distributions in order to estimate their means. To minimize the cumulative regret measure, one would like to sample from the distribution with the highest mean as often as possible (*Exploitation*). At the same time, however, one has to sample all arms at least a little, so as to initially find the best arm (*Exploration*). Any algorithm thus has to handle a trade-off between exploration and exploitation to minimize the regret. How can we organize the sampling in order to achieve this task?

While Gittins (1979) presents an Bayesian-optimal solution to the cumulative regret problem, the Upper-Confidence Bound (UCB) algorithms pioneered by Agrawal (1995) and Auer, Cesa-Bianchi and Fischer (2002) have lead to a blooming literature on multi-armed bandit problems that has been picked up to solve modern problems in website and advertising optimization, where advertisers try to allocate their advertising optimally so as to reduce regret from showing poorly converting advertisements. Several algorithms like the KL-UCB algorithm based on Kullback-Leibler divergence upper confidence intervals (Cappé et al., 2013) or the Bayes-UCB algorithm (Kaufmann et al., 2012a)  have followed and been shown to be optimal as they reach the lower bound introduced in Lai and Robbins (1985) and Burnetas and Katehakis (1996).

@@

Similar algorithms based on indexes can be used on related problems. Instead of minimizing cumulative regret, it can be of interest to find the best arm, that is, the distribution with the largest mean -- with as few samples as possible and with confidence as large as possible. This is particularly useful when there is not a cost from testing arms with lower mean, but when drawing any sample is costly. Thus one needs to explore the distributions as efficiently as possible. Given that the loss function only evaluates the classifiation once at the end of the sampling process, exploitation is *not* necessary, and the setting is thus referred to as *Pure Exploration*. The need for strategies different from uniform sampling arises still: One does not want to continue to sample distributions that one is already confident in. This setting is related to the problem faced in Bayesian Optimization: Training of, for example, deep neural network architectures is costly in terms of time and computation, and so the parameter space needs to be explored with as few samples as possible. (Zhang et al., 2017; Brochu et al., 2010). The problem of best arm identification is also very related to sequential statistical hypothesis testing as noted by Chernoff (1958), and picked up by Garivier and Kaufmann (2016) and Russo (2016), since the experimenter has to make a decision about how the means of distributions relate to each other.

Given the focus on exploration, the pure exploration problem is usually described with a loss function different from the one in the cumulative regret settings. Audibert et al. (2010) introduce a soft measure called *simple regret* based on which they describe UCB policies. Other authors investigate binary measures that are equal to 1 when the best arm has not been identified (Russo, 2016; Garivier and Kaufmann, 2016). These binary measures correspond to the probability of the identified arm being the best, and thus are inherently related to hypothesis testing. The problem of best arm identification can readily be extended to the so called Top-$m$ problem of identifying the best $m$ of $K$ arms (Kaufmann and Kalyanakrishnan, 2013; Kaufmann et al. 2016).

In this thesis, we consider the thresholding bandit problem. While the name is reminiscient of the cumulative regret multi-armed bandit, the problem is more similar to the top-$m$ problem. But instead of identifying the best $m$ out of $K$ arms, one aims to find all distributions out of $K$ that have a mean at least as large as a fixed threshold (it is not known upfront how many distributions fulfill this condition). Given that an implicit threshold arises in the top-$m$ problem between the $m$-th best and the $(m+1)$-th best arm (Kaufmann et al., 2016), the two problems share many characteristics. We discuss this in more detail in \autoref{chap:KullbackLeiblerPureExploration}. In the setting described in Locatelli et al. (2016), algorithms in the thresholding bandit problem are evaluated against a binary loss function penalizing once after the sampling process if an arm has not been classified correctly as being above or below the threshold. This setting thus focuses on exploration similar to algorithms employed for best arm identification. Indeed, Locatelli et al. (2016) show how their thresholding algorithm can be used for best arm identification when the mean of the best arm is known upfront.

@@

The app will display a "Good service" or "This place is trendy" description for the restaurant if $\mu_{\text{service}}$ and $\mu_{\text{trendy}}$ are estimated to be above a certain threshold, for example $\tau = 0.5$. That is, more than half of the user base would describe the place as having good service, or as trendy.

To make the app as useful as possible to other users, this information should be shown as quickly as possible. But even though the app has thousands of users, not that many people visit a specific restaurant, and even less people report this back to the app. Thus the company has only a limited budget of $T$ potential users from which to collect samples $X_{\text{service}}$ and $X_{\text{trendy}}$. How can the company ask these $T$ questions to be as confident as possible in its classification of the restaurant's features, that is, in the information it displays?

A straightforward way of sampling users would be to ask the two questions in a round-robin fashion: Ask for "Service", then for "Trendy", then for "Service", ... Given 100 users, 50 users give feedback on the service level, 50 users report on the trendiness level. This, however, is unlikely to be optimal with respect to the confidence after 100 samples. Given that the company asks the users sequentially, there is the option to adapt the sampling strategy over time: If the first ten users report $\sum_{s=1}^{5}X_{\text{service},s}=5$, while $\sum_{s=1}^{5}X_{\text{trendy},s}=3$, then the company might already be quite confident that $\mu_{\text{service}} \geq \tau = 0.5$, while this is less clear for $\mu_{\text{trendy}}$. Instead of continuously sampling both questions, it might be more useful to collect more feedback on the trendiness level so as to ensure its position relative to the threshold. After all, the company would like to use the 100 samples to maximize its confidence in *both* displayed tips.

Finding the optimal sequential sampling strategy to classify distributions against a threshold with as much confidence as possible after a limited budget of samples is the problem described above. It is a setting described by the thresholding bandit problem, which we discuss for the remainder of this thesis. In the following section, we define setting of the thresholding bandit problem more generally.

@@

Consider a setting with unknown probability distributions $\nu_i$ indexed by $i \in \mathbb{A}=\{1, ..., K\}$. To stick with the idea of multi-armed bandits, we call the distributions arms, and observations from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, it would like to estimate the means $\mu_1, ..., \mu_K$ of each distribution $\nu_i$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from the different arms, where $X_{i} \sim \nu_i$. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which it would like to observe the reward $X_{I_t,t}$. The decision of which arm to pull next is often made by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then $B_i(t)$ is a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indexes can be called strategy (Kaufmann et al., 2012a) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

@@

The thresholding problem for the learner is now to use a budget of overall $T$ samples (or rounds) to explore the different arms. After this exploration phase, the learner returns a set of arms that it believes to be over the threshold. When evaluating the classification of the learner, arms with mean inside of the epsilon interval $\tau \pm \epsilon$ are not considered. The learner only has to correctly distinguish the means from the threshold up to the precision given by $\epsilon$. Thus, the algorithm is not evaluated against $\mathcal{S}_{\tau}$ and $\mathcal{S}^C_{\tau}$, but needs to distinguish $\mathcal{S}_{\tau+\epsilon}$ from $\mathcal{S}^C_{\tau-\epsilon}$.

The learner's classification is returned as the set $\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) \subseteq \mathbb{A}$ after $T$ rounds; that is, the set of arms estimated to be over the threshold. The classification is evaluated against the binary loss given by (Locatelli et al., 2016):

@@

Algorithms for this thresholding problem can be compared based on their expected loss. Let $\mathbb{E}$ be the expectation with regard to the samples $X_{i,1}, ..., X_{i,T}$ collected by the algorithm. Then

@@

is the expected loss, or the probability of making at least one wrong classification. Both the probability of not identifying an arm that is actually over the threshold, as well as the probability of accepting an arm as over the threshold when it is not, are evaluated. The oracle strategy would evaluate all arms correctly with an error probability of $0$ (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit problem setting which we will use throughout the thesis. The setting can be extended by different assumptions to make certain algorithms or bounds feasible. For that reason, Locatelli et al. (2016) assume that the distributions $\nu_i, i \in \mathbb{A}$ are $R$-sub-Gaussian. Under this assumption, the authors are able to prove a lower bound on the largest probability of error of the best algorithm in the setting just introduced, as well as an upper bound on the performance of their APT algorithm which we will show in the following chapter. As in their paper, we define $R$-sub-Gaussian as follows:

@@

As noted by Locatelli et al. (2016), this ensures that the mean $\mu_i$ of the distributions $\nu_i$ considered in our problem are finite. $R$-sub-Gaussian distributions offer concentration inequalities for the concentration of the empirical mean around the true mean, which makes it possible to deduce the bounds presented by the authors and repeated below. Among other distributions, the family of $R$-sub-Gaussian distributions encompasses Gaussian distributions of variance $R^2$ for $R \in \mathbb{R}$, as well as all bounded distributions, with the important case of Bernoulli distributions.

Having described the thresholding bandit setting as well as the assumption of $R$-sub-Gaussianity, we now proceed to describe the *Anytime Parameter-free Thresholding* (APT) algorithm proposed in Locatelli et al. (2016) for the thresholding bandit problem.

@@

\IncMargin{1em}
\begin{algorithm}
\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$}
\Output{$\hat{S}_{\tau} = \{i \in \{1,...,K\}, \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{i \in A} B_i^{APT}(t) = \arg \min_{i \in A} \sqrt{T_i(t-1)} \cdot(|\hat{\mu}_i(t-1) - \tau|+\epsilon)$}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{Anytime Parameter-free Thresholding Algorithm (Locatelli et al., 2016)}\label{alg:APT}
\end{algorithm}
\DecMargin{1em}

Just like the general index policies mentioned in \autoref{sec:Setup}, we see that the APT algorithm adapts its strategy using an index $B_i(t+1)$ for each arm that is based entirely on the observations observed until time $t$. Given the arm with the smallest index is pulled at round $t+1$, we observe that arms estimated to be further away from the threshold $\tau$ as measured by the absolute distance $\hat{\Delta}_i(t)$ to their empirical mean $\hat{\mu}_i$ are pulled less often. Intuitively, it is easier to classify arms as below or above $\tau$ when their mean is far away from the threshold. In contrast to the uniform sampling strategy, APT will gather more evidence for arms which appear to be inherently more difficult to classify. Likewise, APT tends to less often pull arms which it already pulled considerably often as measured by $T_i(t)$. If arm $i$ has been pulled more often than arm $j$, then we should in general be more confident in its estimated mean and the corresponding distance to the threshold. If we are confident in $i$, then it is better to explore other arms by drawing an observation from arm $j$. By adapting its sampling strategy in this way during the sequential sampling process, the APT algorithm minimizes the probability of error $\mathbb{E}[\mathcal{L}(T)]$ after $T$ samples have been observed.

The previous paragraph describes how arms close to the threshold are intuitively more difficult to classify. In the index employed by the APT algorithm, the distance between an arm and the threshold is measured by the absolute distance $\hat{\Delta}_i(t)$ between the (estimated) mean of the arm and the threshold $\tau$ (and the epsilon interval). When the APT algorithm realizes that an arm has a mean close to the threshold, then this arm requires more attention than an arm far away from the threshold; the confidence in the empirical mean has to be larger. Or the empirical mean has to be better concentrated around the true mean. Otherwise it is easy to make a mistake on this arm. To gain this confidence, the arm requires more samples, and thus a larger share of the overall $T$ samples available. With this intuition, it makes sense to measure the complexity of a thresholding problem: If we knew upfront how complex a given set of arms is, we would know whether to expect a small or large probability of error given our budget $T$. Also, for example in the best arm identification problem, there exist algorithms which take the complexity of a problem as input. Based on the complexity they then adapt their sampling strategy. See for example the UCB-E algorithm in Audibert et al. (2010) for the best arm identification problem, or the KL-LUCB-E algorithm in Kaufmann and Kalyanakrishnan (2013) as extension to the top-$m$ problem.

While the APT algorithm does not require the complexity to be known upfront (which would be problematic as it is *not* known upfront), it is necessary to define the complexity to deduce the upper bound on its probability of error. Similar to the distance measure used in the algorithm's index, we define the gap $\Delta_i^{\tau,\epsilon}$ between arm $i$ and the threshold $\tau \pm \epsilon$ as

@@

Note that the complexity of the problem increases in the number of arms $K$ and decreases in the gap $\Delta_i^{\tau,\epsilon}$. Consequently, when we require less precision, that is, we increase $\epsilon$, the hardness decreases. Indeed, by using $\epsilon > 0$, we upper bound the complexity of a problem. Consider a problem with $K$ arms. Then even in the worst case of $\mu_i = \tau$ for all $i \in \mathbb{A}$, $H \geq \frac{K}{\epsilon^2}$. Furthermore, we see that each arm contributes its individual complexity, $h_i^{\tau,\epsilon} = \frac{1}{(\Delta_i^{\tau,\epsilon})^2}$ linearly into the problem. Thus, if the complexity $H$ is a good measure of the actual problem complexity, and we knew upfront which arm has which mean, we could allocate the "optimal" share of the budget, $T \cdot h_i /H$ to arm $i$. However, since we know neither $\mu_i$ nor $H$ upfront, the best we can do is to adapt our allocation of samples more and more towards this optimum as we observe samples.

At this point, one can question whether the measure of complexity $H$ proposed in Locatelli et al. (2016) is a good measure of the "true" problem complexity. In the following \autoref{sec:UpperBoundLocatelli}, we will show the upper bound on the probability of error for the APT algorithm and see how the complexity $H$ arises naturally for the APT algorithm. In \autoref{sec:LowerBoundLocatelli} we then show the lower bound presented in Locatelli et al. (2016), and that the APT algorithm is optimal in that it matches the lower bound. We discuss the results and the APT algorithm further in \autoref{sec:DiscussionLocatelli}.

## Upper Bound for the Anytime Parameter-free Thresholding Algorithm \label{sec:UpperBoundLocatelli}

@@

This bound is non-trivial at much smaller values of $T$, and we will use it in the following.
